Death By Dropdown?

### 2025-04-01 conference-materials/2025_shiny-conf/slides.qmd
Revises slide 6
- Adds a portrait of Isaac Newton and his quote to emphasize the foundation of building upon existing work.
- Updates slide content to highlight benefits to both users and developers, focusing on integration of psychology and visual communication fields.

- Walk through the key issues with the "before" dashboard:
  * "Look at all those filters - users had to make decisions before seeing ANY data"
  * 8 tabs of dense data tables...need to know what you want to figure out in advance
  * "No guidance on where to start - cognitive overload from the beginning"
  * No one wants to use a dashboard like this...and no one will...UIs like this are a major contributor to dashboard rot
:::

@@

::: {.nonincremental}
- Builds on established UX heuristics and theories
- Incorporates data storytelling frameworks
- Systematically integrates cognitive bias mitigation
- Cross-disciplinary synthesis (UX, data viz, behavioral economics)
- Stage-based process rather than heuristic lists
- Specifically tailored for analytics dashboards
:::

@@

- Acknowledge that this builds on existing work - "We're standing on the shoulders of giants"
- Emphasize the unique integration of disciplines - behavioral economics + UX + data viz
- Key difference from other approaches: process-oriented vs. checklist approach
- "What makes BID different is that it's specifically designed for analytics dashboards"
- This is important context before diving into the stages in detail
- Quickly mention some influences: "Edward Tufte's data visualization principles, Nielsen's heuristics, and Kahneman's work on cognitive biases"
:::

>>>>

- Even thought SoundPulse is a composite client...this UI is actually a fair representation of what many of our market research clients have come to expect from their dashboards
  * We didn't just pick the worst dashboard we could find...we again, put a composite together to give you a representative idea of the kinds of challenges we've faced as devs trying to help our clients with the products we deliver
- Walk through the key issues with the "before" dashboard:
  * "Look at all those filters - users had to make decisions before seeing ANY data"
  * 8 tabs of dense data tables...need to know what you want to figure out in advance
  * "No guidance on where to start - cognitive overload from the beginning"
  * No one wants to use a dashboard like this...and no one will...UIs like this are a major contributor to dashboard rot
:::

@@

::: {.d-flex .align-items-center}
![](img/newton-portrait.jpg){width="15%" style="float:left; margin-right:15px;"}

> "If I have seen further, it is by standing on the shoulders of giants."
>
> — Isaac Newton
:::

:::: {.columns}

::: {.column}
- **UX Design**: Beyond heuristic lists to guided process
- **Data Storytelling**: From data dumps to meaningful narratives
:::

::: {.column}
- **Psychological Science**: Evidence-based cognitive principles
- **Visual Communication**: Optimized for information perception
:::

@@

- Acknowledge that this builds on existing work - "We're standing on the shoulders of giants"
- Emphasize the unique integration of fields - psychology, UX, data & visual communication
  * Quickly mention some influences: "Edward Tufte's data visualization principles, Kahneman's work on cognitive biases, etc."
- Key difference from other approaches: process-oriented vs. checklist approach
- "What makes BID different is that it's specifically designed for analytic applications (like dashboards)"
  * The ultimate goal of the BID framework is to helping end users gain clearer insights from dashboards...while giving developers a structured approach to building more effective applications
:::
